---
title: "Final_Project"
author: "Henry Lewis"
date: "18/12/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

The dataset chosen for this assignment is called Craft Beers dataset. 
Description: This dataset contains a list of 2,410 US craft beers and 510 US breweries and its available at https://www.kaggle.com/nickhould/craft-cans

```{r}
# Import Libraries 
library("dplyr")
```

# Part 1: Analysis

## Beer Dataset
```{r}
# Read the beers csv to a variable beer with first row containing column names and removing white space
beer <- read.csv("beers.csv", header = TRUE, strip.white=TRUE) 
# set the column name to beer_id
colnames(beer)[4] <- c("beer_id")
# set the column name to beer_name
colnames(beer)[5] <- c("beer_name")
# Remove the index column
beer_df <- beer[,-1]
```

## Breweries Dataset
```{r}
# Read the breweries csv to a variable breweries with first row containing column names and removing white space
breweries <- read.csv("breweries.csv", header = TRUE, strip.white=TRUE)
# set the column name to brewery_name
colnames(breweries)[2] <- c("brewery_name")
# set the column id to brewery_id
colnames(breweries)[1] <- c("brewery_id")
```

## Merged dataset
```{r}
# merge the beer and breweries and remove the first column
df <- merge(beer_df, breweries, by.x = "brewery_id")
df <- df[,-1]
head(df)
```



### Breweries Analysis
```{r}
breweries_df <- breweries[,-1]
head(breweries_df)
```
The breweries dataframe contains `r nrow(breweries_df)` observations and `r ncol(breweries_df)` columns that include the brewery name, city location, and state within the United States where the brewery is located.

```{r}
# Structure of teh Breweries dataset
str(breweries_df)

```



## Analyse the number of Breweries 


```{r}
#Number of breweries per state
state_breweries <- table(breweries_df$state)
state_breweries
```


```{r, echo=TRUE}
max5_state_breweries <-head(sort(state_breweries,decreasing = TRUE),5)
max5_state_breweries

```


```{r}
colors = c("gray", "orange", "blue")
barplot(max5_state_breweries, main = "Top 5 Number of Breweries by State", xlab ="US State", ylab = "Number of Breweries",
col= colors, las = 1)
```

As can be seen Colorado `r names(max5_state_breweries[1])` has the largest quantity of breweries with `r max(state_breweries)`. Then comes California with `r max5_state_breweries[2]`, Michigan with `r max5_state_breweries[3]`, Oregon with `r max5_state_breweries[4]` and Texas with `r max5_state_breweries[5]`.

```{r}
colorado_brew <- breweries_df[which(breweries_df$state == "CO"),]
colorado_breweries <- colorado_brew[1]
nrow(colorado_breweries)

```

```{r}
colorado_brew_cities <- colorado_brew[1:2]

brewery_cities <- colorado_brew_cities %>%
group_by(city) %>% summarize(n())

brewery_cities <- as.data.frame(brewery_cities)
colnames(brewery_cities)[1] <- c("city")
colnames(brewery_cities)[2] <- c("brew_num")

```

```{r}
library(ggplot2)
ggplot(brewery_cities, aes(x=city, y=brew_num, size = brew_num)) + 
    geom_point(color = "darkgreen")+ 
  # angles the labels
  theme(axis.text.x = element_text(angle = 45, hjust=1))+ 
  # truncates labels 
  scale_x_discrete(label = function(x) stringr::str_trunc(x, 13)) + ggtitle("Number of Breweries in Colorado Cities") + 
  # Adds a theme and adds a centered title
  theme(plot.title = element_text(hjust = 0.5))+
  # Adds detailed caption information alongside a theme
  labs(caption = "Data source: kaggle",
              x = "Cities in Colorodo", y = "Number of Breweries") + theme(
                                            panel.background = element_rect(fill = "lightblue",
                                                                          colour = "lightblue",
                                                                          size = 0.5, linetype = "solid"),
                                            panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                                                          colour = "white"), 
                                            panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                                                          colour = "white")
                                            )

```

## Analyse the Beers
```{r}
# first 5 lines of the beers dataset
head(beer_df)
```
The total number of beers contained within the dataset is `r nrow(beer_df)`.  

```{r}
# function for calculating the number of N/A values
beer_missing <- sapply(beer_df, function(x)sum(is.na(x)))

```
The total number of missing values contained within the ABV (Alcohol By Volume) column is `r beer_missing[1] ` representing `r beer_missing[1]/nrow(beer_df)*100`% of the dataset, whilst the number of missing values contained within IBU (International Bitterness Units) is `r beer_missing[2] ` representing `r beer_missing[2]/nrow(beer_df)*100` % of the dataset. The remaining columns have no missing values.

```{r}
beer_averages <- sapply(beer_df[1:2], function(x) mean(x, na.rm = TRUE))
beer_averages
```

```{r}
hist(beer_df$abv)
```

```{r}
hist(beer_df$ibu)
```

```{r fig.height = 5, fig.width = 6, fig.align = "center"}
# count the number of each beer type
beer_type <- beer_df %>% 
  count(style)

df_bt <-beer_type[order(beer_type$n,decreasing = TRUE),]
# get the top 5 beers
top_5_beer_types <- head(df_bt,5)

# Plot the 5 top beer types in the dataset

ggplot(data = top_5_beer_types, aes(x=style, y=n)) + 
    geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+ 
  # truncates labels 
  scale_x_discrete(label = function(x) stringr::str_trunc(x, 20)) + ggtitle("Top 5 Styles of Beer that Appear in Dataset") + 
  # Adds a theme and adds a centered title
  theme(plot.title = element_text(hjust = 0.5))+
  # Adds detailed caption information alongside a theme
  labs(x = "Top 5 Beer Styles in Dataset", y = "Number of Beers")
```

- The top beer present in the dataset is American IPA followed by America Pale Ale. The American IPA is almost twice as popular as the America pale ale. The other beers are represented in fewer quantities within the dataset than the top two beers.

```{r}
# Beer with the highest abv
abv <- beer_df[order(beer_df$abv,decreasing = TRUE),]
ibu <- beer_df[order(beer_df$ibu,decreasing = TRUE),]

max_abv <- head(abv,1)
min_abv <- tail(abv,1) 

max_ibu <- head(ibu,1)
min_ibu <- tail(ibu,1)
max_ibu
```
- The beer with the highest ABV is `r max_abv['beer_name']` with an ABV of `r max(beer_df$abv,na.rm =   TRUE)*100` %, whilst the beer with the lowest ABI is `r min_abv['beer_name']` with an ABV of `r min(beer_df$abv,na.rm = TRUE)*100` %. 
- The beer with the highest IBU is `r max_ibu['beer_name']` with an ABI of `r max(beer_df$ibu,na.rm = TRUE)`, whilst the beer with the lowest ABI is `r min_ibu['beer_name']` with an ABI of `r min(beer_df$ibu,na.rm = TRUE)`.


```{r}
# graph of ABV vs IBU
x <- beer_df$abv
y <- beer_df$ibu
ggplot(data = beer_df, aes(x=ibu, y=abv)) + 
    geom_point( color="#69b3a2")+
  geom_smooth(method=lm , color="red", se=FALSE) 
  
```

# Part 2: R Package

As part of this section, I have chosen to explore the equisse package. This is an R package that facilitates data manipulation for exploratory data analysis. It is one of the core packages of the tidyverse. The main goal of this package is to make it easier to manipulate data and allows the operators to easily extract, rearrange and manipulate to provide insights into the datasets in a user friendly way. The package also allows the use of pipes that allow the output of one command to become the input of another command.  As part of this analysis, the previous datasets will utilized as well as the combined dataset of beers and breweries. 

```{r}
library(esquisse)
```

```{r}
# getting the highest 5 average abv 
highest_avg_abv <- beer_df %>% 
                  group_by(style) %>% 
                  summarize(Mean = mean(abv, na.rm=TRUE))
highest_average_abv <- highest_avg_abv[order(highest_avg_abv$Mean,decreasing = TRUE),]
high_5_abv <- head(highest_average_abv,5)
```

```{r}
#install.packages("DataExplorer")
#create_report(high_5_abv)
```


```{r}
# getting the highest 5 average ibu
highest_avg_ibu <- beer_df %>% 
                  group_by(style) %>% 
                  summarize(Mean = mean(ibu, na.rm=TRUE))
highest_average_ibu <- highest_avg_ibu[order(highest_avg_ibu$Mean,decreasing = TRUE),]
head(highest_average_ibu,5)
```

```{r}
# getting the lowest 5 average ibu
min_ibu <- beer_df %>% 
                  group_by(style) %>% 
                  summarize(Mean = mean(ibu, na.rm=TRUE))

mimimum_ibu <- min_ibu[order(min_ibu$Mean,decreasing = FALSE),]
head(mimimum_ibu,5)
```

```{r}
# getting the state with the highest 10 average ibu
avg_state_ibu <- df %>% 
                group_by(state)%>%
                summarize(Mean = mean(ibu, na.rm=TRUE))
average_state_ibu <- avg_state_ibu[order(avg_state_ibu$Mean,decreasing = TRUE),]
average_state_ibu <- head(average_state_ibu,10)


# Plot the states with the highest 10 average ibu in the dataset

ggplot(data = average_state_ibu, aes(x=state, y=Mean, fill=Mean)) + 
    geom_bar(stat="identity", color="black", position=position_dodge())+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+
  # truncates labels 
  scale_x_discrete(label = function(x) stringr::str_trunc(x, 20)) + ggtitle("States with highest in Dataset") + 
  # Adds a theme and adds a centered title
  theme(plot.title = element_text(hjust = 0.5))+
  # Adds detailed caption information alongside a theme
  labs(x = "States with ", y = "Number of Beers")
```


```{r}
# Getting the cities with the highest 10 average ibu
avg_city_ibu <- df %>% 
                group_by(city)%>%
                summarize(Mean_IBU= mean(ibu, na.rm=TRUE))
average_city_ibu <- avg_city_ibu[order(avg_city_ibu$Mean_IBU,decreasing = TRUE),]
average_cities_ibu <- head(average_city_ibu,10)

# Plotting the cities with the highest 10 average ibu
ggplot(data = average_cities_ibu, aes(x=city, y=Mean_IBU, fill=Mean_IBU)) + 
    geom_bar(stat="identity", color="black", position=position_dodge())+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+
  # truncates labels 
  scale_x_discrete(label = function(x) stringr::str_trunc(x, 20)) + ggtitle("Cities with highest IBU") + 
  # Adds a theme and adds a centered title
  theme(plot.title = element_text(hjust = 0.5))+
  # Adds detailed caption information alongside a theme
  labs(x = "Cities", y = "Mean IBU")
```

# Part 3: Functions/Programming
The third part of the  assignment 

```{r}
stateAnalysis <- function(state){
  
  # Creating a 
  StateParameters <- list(name=state, 
              beer_mean_ibu_state = mean(df[,"ibu"][df$state == state],na.rm=TRUE ),
              beer_mean_abv_state = mean(df[,"abv"][df$state == state], na.rm=TRUE),
              beer_max_ibu_state = max(df[,"ibu"][df$state == state],na.rm=TRUE ),
              beer_max_abv_state = max(df[,"abv"][df$state == state],na.rm=TRUE ),
              beer_min_ibu_state = min(df[,"ibu"][df$state == state],na.rm=TRUE ),
              beer_min_abv_state = min(df[,"abv"][df$state == state], na.rm=TRUE)
              )
  class(StateParameters) <- "state"
  return(StateParameters)
}
stateAnalysis("CO")
```



```{r}
summary.Region <- function(obj){ 
  cat("The mean values of region", obj$name, "are as follows:\n")
  cat("Mean Beer per capita: ", obj$beer_mean, "|", "stdev: ", obj$beer_sd, "\n")
  cat("Mean Wine per capita:", obj$wine_mean, "|", "stdev: ", obj$wine_sd, "\n")
  cat("Mean Spirit per capita:", obj$spirit_mean, "|", "stdev: ", obj$spirit_sd,  "\n")

}
```

```{r}
cityAnalysis <- function(city){
  
  # Creating a 
  CityParameters <- list(name=city, 
              beer_mean_ibu_city = mean(df[,"ibu"][df$city == city],na.rm=TRUE),
              beer_mean_abv_city = mean(df[,"abv"][df$city == city], na.rm=TRUE),
              beer_max_ibu_city = max(df[,"ibu"][df$city == city],na.rm=TRUE ),
              beer_max_abv_city = max(df[,"abv"][df$city == city],na.rm=TRUE ),
              beer_min_ibu_city = min(df[,"ibu"][df$city == city],na.rm=TRUE ),
              beer_min_abv_city = min(df[,"abv"][df$city == city], na.rm=TRUE)
              )
  class(CityParameters) <- "city"
  return(CityParameters)
}

cityAnalysis("Louisville")
```


```{r}
BeerAnalysis <- function(style){
  
  # Creating a 
  BeerParameters <- list(name=style, 
              mean_ibu_for_beer_style = mean(df[,"ibu"][df$style == style],na.rm=TRUE),
              mean_abv_style_for_beer_style = mean(df[,"abv"][df$style == style],na.rm=TRUE)
              
              )
  class(BeerParameters) <- "beer"
  return(BeerParameters)
}

BeerAnalysis("American Pale Lager")
```

